Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-02-29 paper/ms.tex
additional comp in rad core + revised discussion

Given that magnetic inhibition of convection is required to explain the anomalously young age of low-mass K- and M-type stars, it is worth considering at which mass or effective temperature the effects related to magnetism cease to play an important role in governing stellar structure. 
Looking at Fig.\ \ref{fig:ages}a, early- to mid-G stars follow a 10 Myr standard stellar model isochrone. Instead of continuing the trend, late-G stars appear to slope toward a 5 Myr isochrone around \teff\ $\sim 5\,750$ K. The empirical sequence exhibits a shallower slope than the model isochrone at 10 Myr.
 %, evidenced by a flattening of the empirical locus.

The shallower slope may be an artifact of how the empirical locus was computed, {\bf but this seems unlikely to be a dominant factor as the HR diagram is well-populated at this effective temperature. It is somewhat under-populated in the immediate vicinity of \teff~$= 5\,500$~K, but is once again well-populated near 4\,500~K. The general trend should therefore be robust against errors in how the empirical locus is computed.}

{\bf Errors transforming from photometric colors to effective temperatures may also provide a plausible explanation for the shallower slope of the HR diagram in the vicinity of 5\,000~K. Below this temperature, stellar atmospheric opacity becomes increasingly dominated by molecular species, particularly TiO in the optical and H$_2$O in the infrared, which poses problems when attempting to derive a robust effective temperature scale. If the effective temperature scale adopted by \citet{Preibisch1999} is too cool by about 200 K, the observed transition might disappear and low-mass K- and M-type stars would fall closer to the 10 Myr standard model isochrone.

\citet{Preibisch1999} constructed an spectral type to effective temperature transformation at a gravity intermediate between dwarf and giant scales using the transformation derived for luminosity class IV objects by \citet{deJager1987}. This temperature scale is consistent with the more recent scale of \citet{Herczeg2015} who used PHOENIX BT-Settl model atmospheres \citep{Allard2011} to infer effective temperatures from photometric colors. Agreement between empirical methods and theoretical predictions is reassuring. However,
comparing these results to the 5 -- 30 Myr pre-main-sequence star temperature scale proposed by \citet{Pecaut2013}, one finds that the temperature scale adopted by \citeauthor{Preibisch1999} is about 100~K cooler for early K-type stars, but the two temperature scales agree around spectral type M0. For early-to-late M-type stars, \citeauthor{Preibisch1999} find temperatures about 100~K warmer than \citet{Pecaut2013}, a reversal from the comparison for late-K-type stars. Considering these differences, the sharp knee feature in the HR diagram may be accentuated by the adoption of an erroneous temperature scale, but it does not appear sufficient to fully explain the shallower slope observed in the HR diagram.}

Instead of being the result of errors in the data, the shallower slope in the HR diagram may indicate {\bf a real transition} where the influence of magnetic inhibition of convection becomes {\bf a more critical component governing a star's structure}.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.95\linewidth]{radiative_core_masses.eps}
    \caption{Mass fraction of the stellar radiative core as a function of (a) total stellar mass and (b) effective temperature. }
    \label{fig:rad_core}
\end{figure}

{\bf Predictions from stellar evolution models suggest a transition should exist. Convective envelopes in stars with masses greater than $\sim1.2$ $M_{\odot}$ gradually recede from the deep interior toward the stellar surface in response to the ignition of the proton-proton ($p$-$p$) chain and continued release of gravitational potential energy as the star contracts. Eventually, the convective envelope in these stars will disappear entirely following ignition of the CN-cycle. As the convective envelope grows thinner, magnetic fields of a given strength will have a larger impact on the thermal structure of the super-adiabatic region. However, a restructuring of the super-adiabatic layers in a star with a thin convective envelope has less of an impact on the thermal structure of the deeper layers, which are assumed to be in radiative equilibrium, than for a star whose interior is largely convective. The thinning of the convective envelope is also accompanied by a drop in surface gas pressure as the star's radius inflates, lowering the permitted equipartition magnetic field strength. The combination of these two effects suggests that the transition from a magnetic sequence to a non-magnetic sequence should coincide with the disappearance of the convective envelope.} % confirm discussion of super-adiabatic layers

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.95\linewidth]{radiative_core_mass_difference.eps}
    \caption{Difference in radiative core mass between magnetic and standard stellar evolution models as a function of effective temperature. The difference is shown as the radiative core mass of a magnetic model minus the radiative core mass of a standard stellar model at the same effective temperature, ($M_{\rm rad.\ core,\, mag} - M_{\rm rad.\ core,\, std}$).}
    \label{fig:delta_mass}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\linewidth]{radius_bump.eps}
    \caption{Brief period of radius inflation described in Sect.~\ref{sec:radius} for stars with masses of 1.60\msun\ (blue solid), 1.55\msun\ (green short-dashed), 1.50\msun\ (blue dash-dotted), 1.45\msun\ (yellow long-dashed). Mass tracks are computed without a magnetic field (i.e., non-magnetic configuration). Effective temperatures at these masses are characteristic of F stars at 10 Myr.}
    \label{fig:bump}
\end{figure}

@@

>>>>

{\bf Figure~\ref{fig:ages} shows that a 10 Myr magnetic model isochrones traces a 5 Myr standard model isochrone for effective temperatures below about 5\,000~K, but traces a 10 Myr standard model isochrone above 6\,000~K. This suggests that magnetic inhibition plays an important role in dictating the structure of cool low-mass stars, but the structure of warmer high-mass stars are relatively insensitive to the influence of magnetic fields. In between these two regimes, the models predict a that a transition occurs in which stars appear to exhibit an increasing sensitivity to magnetic inhibition of convection as effective temperature decreases.

As was noted in Sect.~\ref{sec:results}, Figs.~\ref{fig:empirical} and \ref{fig:ages}a indicate that this transition occurs over a similar effective temperature domain in the HR diagram of Upper Scorpius. Early- to mid-G stars in Upper Scorpius (\teff$ \sim 6000$~K) are best described by a 10 Myr standard stellar model isochrone. Instead of continuing to follow a 10 Myr isochrone, late-G stars appear to be better represented by an isochrone intermediate between 5 and 10 Myr causing the empirical sequence to exhibit a shallower slope than standard model predictions between $5750 > T_{\rm eff} / {\rm K} > 5000$~K. Agreement between the transition region observed in the empirical data and the one predicted by magnetic stellar models suggests there is a physical explanation for the observed transition, but might there be an alternative explanation?
}

{\bf The shallower slope may be an artifact of how the empirical locus was computed. The empirical HR diagram is somewhat under-populated in the immediate vicinity of \teff~$= 5\,500$~K, raising the possibility that the running median has artificially created a smooth transition between warmer and cooler stars. However, the HR diagram is well-populated at slightly warmer and slightly cooler temperatures, suggesting the two regions must be somehow connected. While the precise morphology of the HR diagram in this region may be subject to alteration with additional data, the general trend of should be robust against errors in how the empirical locus is computed.}

{\bf Errors transforming from photometric colors or spectral types to effective temperatures may provide a plausible explanation for the shallower slope of the HR diagram in the vicinity of 5\,000~K. Below this temperature, stellar atmospheric opacity becomes increasingly dominated by molecular species, particularly TiO in the optical and H$_2$O in the infrared, which poses problems when attempting to derive a robust effective temperature scale. This is particularly problematic for the HR diagram shown in Fig.~\ref{fig:empirical} because stars above \teff$ \sim 5\,000$~K were transformed from the observational to theoretical plane by different authors (\citealt{Pecaut2012} at warmer temperatures and \citealt{Preibisch1999} at cooler temperatures). If the effective temperature scale adopted by \citet{Preibisch1999} for cool stars is too cool by about 200 K, the observed transition might disappear and low-mass K- and M-type stars would fall closer to the 10 Myr standard model isochrone.

\citet{Preibisch1999} constructed an spectral type to effective temperature transformation at a gravity intermediate between dwarf and giant scales using the transformation derived for luminosity class IV objects by \citet{deJager1987}. This temperature scale is consistent with the more recent scale of \citet{Herczeg2015} who used PHOENIX BT-Settl model atmospheres \citep{Allard2011} to infer effective temperatures from photometric colors. Agreement between empirical methods and theoretical predictions is reassuring. However,
comparing these results to the 5 -- 30 Myr pre-main-sequence star temperature scale proposed by \citet{Pecaut2013}, one finds that the temperature scale adopted by \citeauthor{Preibisch1999} is about 100~K cooler for early K-type stars, but the two temperature scales agree around spectral type M0. For early-to-late M-type stars, \citeauthor{Preibisch1999} find temperatures about 100~K warmer than \citet{Pecaut2013}, a reversal from the comparison for late-K-type stars. Considering these differences, the sharp knee feature in the HR diagram may be accentuated by the adoption of an erroneous temperature scale, but it does not appear sufficient to fully explain the shallower slope observed in the HR diagram.
}

{\bf The transition region observed in the HR diagram appears to be the result of a real transition in which the structure of stars becomes increasingly insensitive to magnetic inhibition of convection toward higher temperatures. Comparison of the empirical HR diagram to magnetic stellar models suggests that the transition's effective temperature domain begins somewhere between 4\,500~K and 5\,000~K, consistent with the knee observed in the empirical data at \teff$\sim 4\,500$~K. According to magnetic stellar models, these effective temperatures correspond to stars with masses between 1.26 \msun\ and 1.40 \msun. This is significant because stars in this mass regime are expected to have relatively thin convective envelopes that eventually disappear entirely. 
}

{\bf % Not only for stars with M ~ 1.2... all intermediate mass stars above 0.28 Msun. 
Convective envelopes in stars with masses greater than $\sim1.2$ $M_{\odot}$ gradually recede from the deep interior toward the stellar surface in response to the ignition of the proton-proton ($p$-$p$) chain, continued release of gravitational potential energy as the star contracts, and a decreasing radiative opacity due to rising gas temperatures \citep{Henyey1955}. During this time, the star's radius is predicted to increase. Eventually, the radius expansion is halted and the convective envelope disappears entirely following ignition of the CN-cycle. Prior to ignition of the CN-cycle, a thinning of the convective envelope means magnetic fields of a given strength will have a larger impact on the thermal structure of the super-adiabatic region. However, a restructuring of the super-adiabatic layers in a star with a thin convective envelope has less of an impact on the thermal structure of the deeper underlying layers, which are assumed to be in radiative equilibrium, than for a star whose interior is largely convective. The thinning of the convective envelope is also accompanied by a drop in the surface gas pressure as the star's radius inflates thereby lowering the permitted equipartition magnetic field strength. The combination of these two effects suggests that the transition from a magnetic sequence to a non-magnetic sequence should coincide with the disappearance of the convective envelope.} % confirm discussion of super-adiabatic layers

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=0.42\linewidth]{radiative_core_masses.eps} \qquad
    \includegraphics[width=0.48\linewidth]{radiative_core_mass_difference.eps}
    \caption{Mass fraction of the stellar radiative core as a function of (a) total stellar mass and (b) effective temperature. (c) Difference in radiative core mass between magnetic and standard stellar evolution models as a function of effective temperature. The difference is shown as the radiative core mass of a magnetic model minus the radiative core mass of a standard stellar model at the same effective temperature, ($M_{\rm rad.\ core,\, mag} - M_{\rm rad.\ core,\, std}$), shown for two ages of standard models (5 Myr, 10 Myr).}
    \label{fig:rad_core}
\end{figure*}

@@

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=0.95\linewidth]{radiative_core_mass_difference.eps}
%    \caption{Difference in radiative core mass between magnetic and standard stellar evolution models as a function of effective temperature. The difference is shown as the radiative core mass of a magnetic model minus the radiative core mass of a standard stellar model at the same effective temperature, ($M_{\rm rad.\ core,\, mag} - M_{\rm rad.\ core,\, std}$).}
%    \label{fig:delta_mass}
%\end{figure}

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{radius_bump.eps}
    \caption{Brief period of radius inflation described in Sect.~\ref{sec:radius} for stars with masses of 1.60\msun\ (blue solid), 1.55\msun\ (green short-dashed), 1.50\msun\ (blue dash-dotted), 1.45\msun\ (yellow long-dashed). Mass tracks are computed without a magnetic field (i.e., non-magnetic configuration). Effective temperatures at these masses are characteristic of F stars at 10 Myr.}
    \label{fig:bump}
\end{figure}
